---
title: "bmi.calculator"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{bmi.calculator}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This R package is for calculation of BMI (body mass index), and some more computations based on weight loss goals. The package provides simple answers to a user's weight loss goals in terms of how much weight loss/gain and calorie deficit/increase should be aimed for with a target weight and time frame in mind. The package also provides helpful visualizations about BMI and calorie intake change trajectory leading to the target.

This document introduces you to bmi.calculator’s basic set of tools, and shows you how to use them. 

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Installation

You can install the development version of bmi.calculator from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("UBC-MDS/bmi-calculator-r")
```

## Functions

The package contains the following functions

-   `calculate_bmi`: computes user's BMI based on weight and height. Also creates a visual of the BMI on scale
-   `project_bmi`: computes how much average change in BMI should be achieved per week given a targeted weight and the timeframe to reach the goal. Also presents a visual trajectory for BMI
-   `project_calories`: computes how much average calorie intake is ideal per week given a targeted weight and the timeframe to reach the goal. Also presents a visual trajectory for calories
-   `exercise_plan`: suggests possible exercise plans to achieve the targeted weight. Also gives a graph showing how much exercise of each kind is needed per day

## Example

To use the package, load the package library with following command:

```{r import}
library(bmi.calculator)
```

To use the functions, see examples below:

### Calculate BMI

```{r example1, fig.height=4, fig.width=6}
# Weight 100kg, height 1.85 meters, BMI value returned.
calculate_bmi(100, 1.85, return_graph=FALSE)

# Weight 100kg, height 1.85 meters, BMI graph returned.
calculate_bmi(100, 1.85, return_graph=TRUE)
```

### Compute average BMI change per week

```{r example2, fig.height=4, fig.width=6}
# Weight 100kg, height 1.85 meters, BMI goal 25, 30 days to reach goal, return average BMI change per week. 
project_bmi(100, 1.85, 25, 30, return_graph=FALSE)

# Weight 100kg, height 1.85 meters, BMI goal 25, 30 days to reach goal, return plot of Projected BMI trajectory. 
project_bmi(100, 1.85, 25, 30, return_graph=TRUE)
```

### Compute caloric intake per day based in a target weight

```{r example3, fig.height=4, fig.width=6}
# Weight 100kg, height 1.85 meters, male, 25 years old, moderate exercise 3-5 times a week, ideal weight 75kg, 
# 25 days to reach goal, return caloric intake per day based in a target weight. 
project_calories(100, 1.85, 1, 25, 1.6, 75, 25, return_graph=FALSE)

# Weight 100kg, height 1.85 meters, male, 25 years old, moderate exercise 3-5 times a week, ideal weight 75kg, 
# 25 days to reach goal, return plot of Projected Weight Loss. 
project_calories(100, 1.85, 1, 25, 1.6, 75, 25, return_graph=TRUE)
```

### Create an exercise plan

```{r example4, fig.height=4, fig.width=6}
# Weight 100kg, height 1.83 meters, female, aged 27
# Target weight: 68kg in 30 days
exercise_plan(100, 1.83, 2, 27, 68, 30)

# Weight 100kg, height 1.83 meters, female, aged 27
# Target weight: 68kg in 30 days
# But this time, a graph instead:
exercise_plan(100, 1.83, 2, 27, 68, 30, return_graph=TRUE)
```
